\name: A hierarchical Bayesian model for Supernova Cosmology

### 2018-03-19 papers/methods/paper.tex
Adding Alex's suggestions for intrinsic dispersion.

The first layer of the hierarchy represents the latent population of $m_B$, $x_1$ and $c$. This can be thought of as the `true' values for the supernova, and with correct summary statistics and uncertainty would be normally distributed around the observed values such that 
\begin{equation}
P(\hat{m}_B,\hat{x}_1,\hat{c}|m_B, x_1, c,\cov) = \mathcal{N}(\lbrace \hat{m}_B,\hat{x}_1,\hat{c} \rbrace|\lbrace m_B, x_1, c \rbrace, \cov). \label{eq:pop}
\end{equation}
However, the presence of chromatic smearing in the intrinsic supernova population \citep{Guy2010, Chotard2011} means the summary statistics reported from light curve fitting are not Gaussian. We model the extra dispersion included in the observations from colour smearing by adding extra independent uncertainty on the colour observation. As shown in \citep{Kessler2013}, the extra dispersion shows heavy redshift dependence. As the extra dispersion may arise from underlying supernova physics, rather than Malmquist bias, we decide to incorporate redshift dependence in our extra uncertainty. We thus add $\kappa_0 + \kappa_1 z$ to our observed colour uncertainty (in quadrature), where $\kappa_0$ and $\kappa_1$ are subject to strong Cauchy priors centered on zero and with width $0.002$, to enforce minimum possible extra uncertainty in our model. The value of this prior was determined from running many iterations with differing prior strengths on our simple data set with simulated extra dispersion according to the results of \citet{Kessler2013}. Each survey has its own set of $\kappa$ parameters. Furthermore, we use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the shift in the mean colour measured as a function of redshift, and subtract this shift from our observed data. We allow the amount of shift to be parametrised to account for different intrinsic scatter models, where parameter $s_c$ controlls the fraction of bias subtraction -- 0 for no shift, up to 1 for the maximum shift found in simulations \citep[which corresponds to the intrinsic scatter model from][]{Chotard2011}. Unlike $\kappa_0$ and $\kappa_1$, which are redefined for each survey, $s_c$ represents a universal shift in supernovae colour. We place a uniform prior on $s_c$.

>>>>

The first layer of the hierarchy represents the parameters that describe each supernova. For the moment, let us consider a single supernova, and represent its parameters as $\theta$. The underlying supernova model and intrinsic dispersion model gives the probability of $p(y|\theta)$ for obtaining light curve data $y$. As mentioned earlier, we work with summary statistics -- numbers output from a light curve fitter given the input $y$. The function $f$ describes the action of the fitter $\lbrace \hat{m}_B, \hat{x}_1., \hat{c} \rbrace = f(y)$. For notational simplify, let us denote $\eta \equiv \lbrace m_B, x_1, c \rbrace$. The model prediction for our derived data is
\begin{align}
p(\hat{\eta} | \theta) = \int \delta(\hat{\eta} - f(y)) p(y | \theta) \, dy.
\end{align}
To help confuse things, we often consider models that are parametrised as $\theta = \lbrace m_B, x_1, c \rbrace = \eta$, so that the probability of derived data is expressed as
\begin{align}
p(\hat{\eta}|\eta) = \int \delta(\hat{\eta} - f(y)) p(y|\eta) dy. \label{eq:intsmear}
\end{align}
It is common in analytic supernova cosmology models to take that
\begin{align}
p(\hat{\eta}|\eta) \sim \mathcal{N}(\hat{\eta} | \eta, \mathcal{C}), \label{eq:pop}
\end{align}
where $\mathcal{C}$ is a covariance matrix reported from light curve fitting. However, the presence of chromatic smearing in the intrinsic supernova population \citep{Guy2010, Chotard2011} means the summary statistics reported from light curve fitting are not Gaussian and equation \eqref{eq:pop} is violated. We address this problem with a similar method to previous analytic likelihoods -- including extra dispersion in the reported covariance matrix $\mathcal{C}$.

We model the extra dispersion only in colour, and do so by adding extra independent uncertainty on the colour observation. Fully covariant extra dispersion on $\lbrace m_B, x_1, c \rbrace$ was also tested, but it showed negligible improvement over just colour dispersion and was far more computationally inefficient. As shown in \citep{Kessler2013}, the extra colour dispersion shows heavy redshift dependence. As the extra dispersion may arise from underlying supernova physics, rather than Malmquist bias, we decide to incorporate redshift dependence in our extra uncertainty. We thus add $\kappa_0 + \kappa_1 z$ to our observed colour uncertainty (in quadrature), where $\kappa_0$ and $\kappa_1$ are subject to strong Cauchy priors centered on zero and with width $0.002$, to enforce minimum possible extra uncertainty in our model. The value of this prior was determined from running many iterations with differing prior strengths on our simple data set with simulated extra dispersion according to the results of \citet{Kessler2013}. Each survey has its own set of $\kappa$ parameters. Furthermore, we use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the shift in the mean colour measured as a function of redshift, and subtract this shift from our observed data. We allow the amount of shift to be parametrised to account for different intrinsic scatter models, where parameter $s_c$ controls the fraction of bias subtraction -- 0 for no shift, up to 1 for the maximum shift found in simulations \citep[which corresponds to the intrinsic scatter model from][]{Chotard2011}. Unlike $\kappa_0$ and $\kappa_1$, which are redefined for each survey, $s_c$ represents a universal shift in supernovae colour. We place a uniform prior on $s_c$.
